Parsing Postfix log files

### 2007-09-26 doc/logparser.tex
Improved the critiques of: pflogsumm, Sawmill, Isoqlog, AWStats and Anteater.

    \item [pflogsumm.pl] \textit{pflogsumm.pl is designed to provide an
        over-view of postfix activity, with just enough detail to give the
        administrator a ``heads up'' for potential trouble spots.\/}
        \newline \url{http://jimsun.linxnet.com/postfix_contrib.html}
        \newline (Last checked 2007/08/13.)

    \item [Sawmill Universal Log File Analysis and Reporting] is a general
        purpose commercial product which parses 687 log file formats
        (correct as of 2007/04/15) and produces reports.  I have not
        experimented with it due to its limited data extraction facilities;
        it does have three different sets of data for Postfix (as of
        2007/04/15, one is beta), but they do not appear to be interlinked,
        nor does it save sufficient data for the purposes of this project.
        The source code is available in an obfuscated form only (presumably
        for a fee), and the product is quite expensive, as it requires a
        license for each report which is to be generated. \newline
        \url{http://www.thesawmill.co.uk/formats/postfix.html} \newline
        \url{http://www.thesawmill.co.uk/formats/postfix_ii.html} \newline
        \url{http://www.thesawmill.co.uk/formats/beta_postfix.html}
        \newline (Last checked 2007/08/13.)

@@

        \url{http://www.splunk.com}

    \item [Isoqlog] \textit{Isoqlog is an MTA log analysis program written
        in C. It designed to scan qmail, postfix, sendmail and exim logfile
        and produce usage statistics in HTML format for viewing through a
        browser. It produces Top domains output according to Sender,
        Receiver, Total mails and bytes; it keeps your main domain mail
        statistics with regard to Days Top Domain, Top Users values for per
        day, per month and years.\/}  \newline
        \url{http://www.enderunix.org/isoqlog/} \newline (Last checked
        2007/08/13.)

@@

        AWStats will produce simple graphs for many different services, but
        that restricts it to supporting the Lowest Common Denominator: the
        data it will extract from an MTA log file is:
        \texttt{time2, email, email\_r, host, host\_r, method, url, code,
        and bytesd.} \newline
        \url{http://awstats.sourceforge.net/} \newline
        \url{http://awstats.sourceforge.net/awstats.mail.html} \newline
        \url{http://awstats.sourceforge.net/docs/awstats_faq.html#MAIL}
        \newline (Last checked 2007/08/13.)

    \item [Log analyser --- throughput monitor] This utility tracks the
        number of events which occurred over a particular time and warns if
        the frequency of events passes a certain threshold.  It's designed
        to provide real time alerts when dictionary attacks, mail loops or
        similar problems occur.  \newline
        \url{http://home.uninet.ee/~ragnar/throughput_monitor/} \newline
        (Last checked 2007/08/13.)

@@

        Anteater hasn't been updated since November 2003, and doesn't have
        any English documentation.

>>>>

    \item [Pflogsumm] \textit{pflogsumm is designed to provide an over-view
        of Postfix activity, with just enough detail to give the
        administrator a ``heads up'' for potential trouble spots.\/}

        Pflogsumm produces a report designed for troubleshooting, and does
        not support saving any data; parsing is fixed and concentrates on
        extracting only the data required to produce the report.  Both the
        parsing and reporting are difficult to extend, as it is a
        specialised tool.  It does not attempt to correlate log lines by
        queueid or pid, nor does it need to deal with the complications
        encountered during this project.  Pflogsumm produces a useful
        report, and successfully dealt with all the Postfix logs tested
        with.

        \url{http://jimsun.linxnet.com/postfix_contrib.html}
        \newline (Last checked 2007/08/13.)

    \item [Sawmill Universal Log File Analysis and Reporting] is a general
        purpose commercial product which parses 687 log file formats
        (correct as of 2007/04/15) and produces reports.  Its data
        extraction facilities are quite limited, though it does extract
        three different sets of data for Postfix (as of 2007/04/15, one is
        beta), but they do not appear to be interlinked, nor does it save
        sufficient data for the purposes of this project.  Not attempt is
        made to correlate log lines or deal with the difficulties
        documented in sections~\ref{complications} and~\ref{additional
        complications}.\footnote{If any attempt is made there is no
        reference to it in the documentation available on the website.}
        The source code is available in an obfuscated form only (presumably
        for a fee), and the product is quite expensive, as it requires a
        license for each report which is to be generated.  The web
        interface allows creation of dynamic reports based on any field,
        but due to the associated cost the author has not experimented with
        it.

        \url{http://www.thesawmill.co.uk/formats/postfix.html} \newline
        Fields extracted: from, to, server, UID, relay, status, number of
        recipients, origin hostname, origin IP and virus.  The fields
        \textit{server}, \textit{uid\/} and \textit{virus\/} are not
        explained in their documentation: \textit{server\/} is probably the
        server the mail is delivered to, and \textit{uid\/} might be the
        uid of the user submitting mail locally.  Postfix does not perform
        any form of virus checking (though it has many options for
        cooperating with an external virus scanner), so the
        \textit{virus\/} field is a mystery.

        \url{http://www.thesawmill.co.uk/formats/postfix_ii.html} \newline
        Fields extracted: from, to, RBL list, client hostname and client
        IP\@.

        \url{http://www.thesawmill.co.uk/formats/beta_postfix.html}
        \newline
        Fields extracted: from, to, client hostname, client IP, relay
        hostname, relay IP, status, response code, RBL list and message id.

        (Last checked 2007/09/26.)

@@

        \url{http://www.splunk.com/} \newline (Last checked 2007/09/25.)

    \item [Isoqlog] \textit{Isoqlog is an MTA log analysis program written
        in C. It designed to scan qmail, postfix, sendmail and exim logfile
        and produce usage statistics in HTML format for viewing through a
        browser. It produces Top domains output according to Sender,
        Receiver, Total mails and bytes; it keeps your main domain mail
        statistics with regard to Days Top Domain, Top Users values for per
        day, per month and years.\/}

        Due to the paucity of data gathered by this tool the author has not
        experimented with it, but based on the documentation it is obvious
        that it doesn't attempt to gather the breadth of data required for
        in-depth analysis, and presumably doesn't attempt to correlate log
        lines per-connection or per-mail.

        \url{http://www.enderunix.org/isoqlog/} \newline (Last checked
        2007/08/13.)

@@

        AWStats will produce simple graphs for many different services, but
        that restricts it to supporting the Lowest Common Denominator.  The
        data it will extract from an MTA log file is limited to: \newline
        \hspace*{2em} time2, email, email\_r, host, host\_r, method, url,
        code and bytesd.  \newline  There does not seem to be an
        explanation of those fields in the documentation.  As AWStats is a
        generic log processor it can safely be assumed that it has
        no special purpose code to deal with the intricacies of Postfix
        logs.

        \url{http://awstats.sourceforge.net/} \newline
        \url{http://awstats.sourceforge.net/awstats.mail.html} \newline
        \url{http://awstats.sourceforge.net/docs/awstats_faq.html#MAIL}
        \newline (Last checked 2007/08/13.)

    \item [Log analyser --- throughput monitor] This utility tracks the
        number of events which occurred over a particular time and warns if
        the frequency of events passes a certain threshold.  It's designed
        to provide real time alerts when dictionary attacks, mail loops or
        similar problems occur. It doesn't attempt to extract or save data,
        correlate log lines, or any of the more advanced tasks described in
        this document, because it is not designed to do so.

        \url{http://home.uninet.ee/~ragnar/throughput_monitor/} \newline
        (Last checked 2007/08/13.)

@@

        Anteater hasn't been updated since November 2003, and doesn't have
        any English documentation so it's difficult, nigh impossible, to
        accurately comment on what analysis is performs.  It did not run
        successfully when tested, and its parsing would certainly be out of
        date as Postfix has evolved considerably since this tool was last
        updated.  As it neither ran successfully nor has documentation the
        author can read it is difficult to comment on the features it
        provides.
